\documentclass{slides}
\usepackage[utf8]{inputenc}
\usepackage{graphics}
\usepackage{portland}
\usepackage{epsfig}
\usepackage{alltt}
\usepackage{moreverb}
\usepackage{url}
\usepackage{eurosym}
\usepackage[dvips,usenames]{color}

\definecolor{MyLightMagenta}{rgb}{1,0.7,1}
\definecolor{darkgreen}{rgb}{0.1,0.7,0.1}

\newcommand{\darkgreen}[1]{\textcolor{darkgreen}{#1}}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\thistle}[1]{\textcolor{Thistle}{#1}}
\newcommand{\apricot}[1]{\textcolor{Apricot}{#1}}
\newcommand{\melon}[1]{\textcolor{Melon}{#1}}
\newcommand{\dandelion}[1]{\textcolor{Dandelion}{#1}}
\newcommand{\green}[1]{\textcolor{OliveGreen}{#1}}
\newcommand{\lavender}[1]{\textcolor{Lavender}{#1}}
\newcommand{\mylightmagenta}[1]{\textcolor{MyLightMagenta}{#1}}
\newcommand{\blue}[1]{\textcolor{RoyalBlue}{#1}}
\newcommand{\darkorchid}[1]{\textcolor{DarkOrchid}{#1}}
\newcommand{\orchid}[1]{\textcolor{Orchid}{#1}}
\newcommand{\brickred}[1]{\textcolor{BrickRed}{#1}}
\newcommand{\peach}[1]{\textcolor{Peach}{#1}}
\newcommand{\bittersweet}[1]{\textcolor{Bittersweet}{#1}}
\newcommand{\salmon}[1]{\textcolor{Salmon}{#1}}
\newcommand{\yelloworange}[1]{\textcolor{YellowOrange}{#1}}
\newcommand{\periwinkle}[1]{\textcolor{Periwinkle}{#1}}

\newcommand{\names}[1]{\periwinkle{#1}}
\newcommand{\motcle}[1]{\mylightmagenta{#1}}
\newcommand{\classname}[1]{\darkgreen{#1}}
\newcommand{\str}[1]{\yelloworange{#1}}
\newcommand{\defun}[1]{\orchid{#1}}
\newcommand{\ti}[1]{\begin{center}\Large{\textcolor{blue}{#1}}\end{center}}
\newcommand{\alert}[1]{\thistle{#1}}
\newcommand{\lispprint}[1]{\dandelion{#1}}
\newcommand{\lispvalue}[1]{\red{#1}}
\newcommand{\tr}[1]{\texttt{\red{#1}}}
\newcommand{\emc}[1]{\red{#1}}
\newcommand{\lispobj}[1]{\green{\texttt{#1}}}
\def\prompt{{\textcolor{Orchid}{CL-USER>}}}
\newcommand{\promptp}[1]{\textcolor{Orchid}{#1>}}

\newcommand{\Comment}[1]{
\begin{center}
\textcolor{yellow}
{#1}
\end{center}
}

\def\bs{$\backslash$}
\def\inputfig#1{\input #1}
\def\inputtex#1{\input #1}

\begin{document}
\landscape
\setlength{\oddsidemargin}{1cm}
\setlength{\evensidemargin}{1cm}
\setlength{\marginparwidth}{1cm}
\setlength{\parskip}{0.5cm}
\setlength{\parindent}{0cm}
%-----------------------------------------------------------
\begin{slide}\ti{Psychology and software development}
\vskip 0.5cm
\begin{center}
Robert Strandh \\
Université de Bordeaux \\
Bordeaux, France
\end{center}
\vfill\end{slide}
%-----------------------------------------------------------
\begin{slide}\ti{Overview of talk}

\vfill\end{slide}
%-----------------------------------------------------------
\begin{slide}\ti{Two different mind sets}

According to the work of Carol Dweck, people have one of two different
\emph{mind sets}:

\begin{itemize}
\item Fixed, or
\item Growth
\end{itemize}

\vfill\end{slide}
%-----------------------------------------------------------
\begin{slide}\ti{Growth}

  \begin{itemize}
  \item Learning new things comes naturally.
  \item Failure is normal and part of the learning process.
  \item Takes pleasure in the learning process itself.
  \end{itemize}

In other words, this is your stereotypical researcher personality.

\vfill\end{slide}
%-----------------------------------------------------------
\begin{slide}\ti{Fixed}

  \begin{itemize}
  \item Needs to be perfect all the time without effort.
  \item Does not like to learn new things.
  \item Learning is seen as detrimental to performance.
  \item Failure means defective.
  \end{itemize}

\vfill\end{slide}
%-----------------------------------------------------------
\begin{slide}\ti{Mind set and software development}

In a team, a person with a fixed mind set is often well liked,
even admired.

He or she is fast and competent in getting things done.

\vfill\end{slide}
%-----------------------------------------------------------
\begin{slide}\ti{Mind set and software development}

But when things might change (new tools, new methods, new programming
language), the person with a fixed mind set feels threatened.

Because, if things change, that person may have to learn something new
which he or she is not comfortable with.

\vfill\end{slide}
%-----------------------------------------------------------
\begin{slide}\ti{Mind set and software development}

As a result, instead of accepting the changes, the person with fixed
mind set tries to make sure the changes do not happen.

The person will not hesitate using methods such as lying about the
nature of the changes, or ridiculing the people in favor of the
changes.

Since the person is often admired, these methods will frequently be
effective.

\vfill\end{slide}
%-----------------------------------------------------------
\begin{slide}\ti{Example of different mind sets}

Let's say John and Mary are part of a development team.  John has a
fixed mind set, whereas Mary has a growth mind set.

Mary might suggest that the programming language Haskell be used for
the next project, because it seems like a good fit.

John might say:
\begin{itemize}
\item that Haskell is slow,
\item that automatic memory management is not adapted to this project,
\item that Haskell will take too long to learn which will increase the
  cost of the project,
\item that even the people that created Haskell are now abandoning it,
  and even something like
\item that Haskell is only for sissies.
\end{itemize}

\vfill\end{slide}
%-----------------------------------------------------------
\begin{slide}\ti{System 1 and system 2}

Daniel Kahneman discusses two different \emph{systems} in the brain:

\begin{itemize}
\item System 1 makes decisions very quickly without having to do any
  complicated calculations,
\item System 2 is invoked when system 1 is not able to make a
  decision, and it checks whether the decision of system 1 was
  reasonable.
\end{itemize}

You can read about it more in Kahneman's book: Thinking, fast and
slow.

\vfill\end{slide}
%-----------------------------------------------------------
\begin{slide}\ti{System 1 and system 2 example}

On example in Kahneman's book is this:

Two items A and B together cost 1\euro{}.  Item B costs 0.9\euro{}
more than item A.  How much does item A cost?

\vfill\end{slide}
%-----------------------------------------------------------
\begin{slide}\ti{System 1 and Kaizen}

Example:  You have the following code:

\begin{verbatim}
int
fun(int x)
{
  int y = fun2(x, 3);
  return y * x;
}
\end{verbatim}

And your cursor is positioned on the \texttt{return} statement.  You
want to go to the beginning of the definition of the function
\texttt{fun}.

\vfill\end{slide}
%-----------------------------------------------------------
\begin{slide}\ti{System 1 and Kaizen}

You have a choice between:

\begin{itemize}
\item Looking in the manual for your text editor in order to find the
  command that will go to the beginning of a definition.
\item Using arrow keys (or equivalent) or the mouse to go there.
\end{itemize}

System 1 will always tell you to use the second option, even though
the first option will pay off within a few days.

\vfill\end{slide}
%-----------------------------------------------------------
\begin{slide}\ti{Questions?}

Thank you for listening!

Do you have any questions?

\vfill\end{slide}
%-----------------------------------------------------------
%% \begin{slide}\ti{}

%% \vfill\end{slide}
%-----------------------------------------------------------
%% \begin{slide}\ti{}

%% \vfill\end{slide}
%% %-----------------------------------------------------------
%% \begin{slide}\ti{}

%% \vfill\end{slide}
%--------------------------------

\end{document}
 
